Practical Parametricity for GADTs

### 2020-09-29 20-gadts/inf-rules.tex
adding well-definedness of term interps to inf.rules

@@

@@

\begin{dfn}\label{def:set-sem}
The {\em object interpretation} $\oCPOsem{\cdot} : \F \to [\oCPOenv,
  \oCPO]$ is defined by
\begin{align*}
  \oCPOsem{\Gamma;\Phi \vdash \zerot}\rho &= 0\\
  \oCPOsem{\Gamma;\Phi \vdash \onet}\rho &= 1\\
  \oCPOsem{\Gamma; \emptyset
    \vdash \Nat^\Phi
    \,F\,G}\rho &= \{\eta : \lambda \ol{K}. \,\oCPOsem{\Gamma;
    \Phi \vdash
    F}\rho[\Phi := \ol K] 
      \Rightarrow \lambda \ol{K}.\,\oCPOsem{\Gamma; 
        \Phi \vdash G}\rho[\Phi := \ol K] \\ 
      &\hspace{0.3in}|~\forall \ol{K = (K^1, K^2, K^*) : \oCPORT_k}.\\
      &\hspace{0.4in}(\eta_{\overline{K^1}}, \eta_{\overline{K^2}})
      : \oCPORsem{\Gamma; \Phi \vdash F}\Eq_{\rho}[\Phi := \ol K]
      \rightarrow \oCPORsem{\Gamma; \Phi \vdash
        G}\Eq_{\rho}[\Phi := \ol K] \} \\
  \oCPOsem{\Gamma;\Phi \vdash \phi\ol{F}}\rho &=
  (\rho\phi)\,\ol{\oCPOsem{\Gamma;\Phi \vdash
    F}\rho}\\
  \oCPOsem{\Gamma;\Phi \vdash F+G}\rho &=
  \oCPOsem{\Gamma;\Phi \vdash F}\rho +
  \oCPOsem{\Gamma;\Phi \vdash G}\rho\\
  \oCPOsem{\Gamma;\Phi \vdash F \times G}\rho &=
  \oCPOsem{\Gamma;\Phi \vdash F}\rho \times
  \oCPOsem{\Gamma;\Phi \vdash G}\rho\\ 
  \oCPOsem{\Gamma;\Phi,\ol\gamma \vdash (\mu \phi.\lambda
    \ol{\alpha}. H)\ol{G}}\rho &= (\mu
    T^\oCPO_{H,\rho})\ol{\oCPOsem{\Gamma;\Phi,\ol\gamma \vdash G}\rho}\\
    \text{where } T^\oCPO_{H,\rho}\,F & = \lambda
  \ol{A}. \oCPOsem{\Gamma;\ol\gamma,\phi, \ol{\alpha} \vdash
    H}\rho[\phi :=  F][\ol{\alpha := A}]\\
  \text{and } T^\oCPO_{H,\rho}\,\eta &= \lambda
  \ol{A}. \oCPOsem{\Gamma;\ol\gamma,\phi, \ol{\alpha} \vdash
    H}\id_\rho[\phi := \eta][\ol{\alpha := \id_{A}}] \\
  \oCPOsem{\Gamma;\Phi \vdash (\Lan^{\ol{\alpha}}_{\ol{K}} F)\ol{A}}\rho &=
    \{ t : (\textit{Lan}_{\ol{\oCPOsem{\Gamma; \ol\alpha \vdash
          K}\rho[\ol{\alpha := \_}]}} \oCPOsem{\Gamma; \Phi, \ol\alpha
      \vdash F}\rho[\ol{\alpha := \_}]) \ol{\oCPOsem{\Gamma; \Phi
        \vdash A}\rho}\\
    &\hspace{0.3in}|~(t,t) \in \oCPORsem{\Gamma;\Phi \vdash
      (\Lan^{\ol{\alpha}}_{\ol{K}} F)\ol{A}}\Eq_\rho\}\\
\end{align*}
\end{dfn}
\new{WHERE DOES THIS ALL GO? Must explain $\Lan$s, say why they are
  functors, say they can be computed using colimits in $\set$ and
  $\rel$, define injections, etc. If $f : B \to B'$ in $\set$, then
  the functorial action of $\mathit{Lan}_{\ol K}\,F$ on $f$ is the
  unique morphism $(\mathit{Lan}_{\ol K}\,F)f$ such that
  \begin{equation}\label{def:functorial-action-lan}
(\mathit{Lan}_{\ol K}\,F)f \circ \iota_{X,g} = \iota'_{X, f \circ
      g}~~~MISSING~SOME~BARS
\end{equation}
Here, $\iota_{X,g}$ is the injection of $FX$ into $(\mathit{Lan}_{\ol
  K}\,F)\,B = \colim{X : \set_0,\, g : KX \to B}{FX}$ and
$\iota'_{X,g'}$ is the injection of $FX$ into $(\mathit{Lan}_{\ol
  K}\,F)\,B' = \colim{X : \set_0,\, g' : KX \to B'}{FX}$.

@@

**Say somewhere that recognizing the non-parallel nature of the Lan
interps, and yet still being able to define the set and rel interps so
that everything is intertwined in exactly the right way to make the
proofs (e.g., of the IEL) work, is what's hard here. This is where the
insight is required.**}

@@

>>>>

\subsection{Preliminaries}

\new{WHERE DOES THIS ALL GO? Must explain $\Lan$s, say why they are
  functors, say they can be computed using colimits in $\set$ and
  $\rel$, define injections, etc. If $f : B \to B'$ in $\set$, then
  the functorial action of $\mathit{Lan}_{\ol K}\,F$ on $f$ is the
  unique morphism $(\mathit{Lan}_{\ol K}\,F)f$ such that
  \begin{equation}\label{def:functorial-action-lan}
(\mathit{Lan}_{\ol K}\,F)f \circ \iota_{X,g} = \iota'_{X, f \circ
      g}~~~MISSING~SOME~BARS
\end{equation}
Here, $\iota_{X,g}$ is the injection of $FX$ into $(\mathit{Lan}_{\ol
  K}\,F)\,B = \colim{X : \set_0,\, g : KX \to B}{FX}$ and
$\iota'_{X,g'}$ is the injection of $FX$ into $(\mathit{Lan}_{\ol
  K}\,F)\,B' = \colim{X : \set_0,\, g' : KX \to B'}{FX}$.
**Say somewhere that recognizing the non-parallel nature of the Lan
interps, and yet still being able to define the set and rel interps so
that everything is intertwined in exactly the right way to make the
proofs (e.g., of the IEL) work, is what's hard here. This is where the
insight is required.**}

@@

{\color{red} We need to restrict set and relational environments to
  $\set_0$ and $\rel_0$ for interps of $\Lan$-types and -terms, and
  note that this is OK because these are our presentable objects, all
  (other) objects can be written as $\omega$-directed colimits of
  presentables, and our interpretations are $\omega$-cocontinuous and
  therefore preserve $\omega$-directed colimits. Say once here, and
  then just use throughout.}

@@

\begin{dfn}\label{def:set-sem}
The {\em object interpretation} $\oCPOsem{\cdot} : \F \to [\oCPOenv,
  \oCPO]$ is defined by
\begin{align*}
  \oCPOsem{\Gamma;\Phi \vdash \zerot}\rho &= 0\\
  \oCPOsem{\Gamma;\Phi \vdash \onet}\rho &= 1\\
  \oCPOsem{\Gamma; \emptyset
    \vdash \Nat^\Phi
    \,F\,G}\rho &= \{\eta : \lambda \ol{K}. \,\oCPOsem{\Gamma;
    \Phi \vdash
    F}\rho[\Phi := \ol K] 
      \Rightarrow \lambda \ol{K}.\,\oCPOsem{\Gamma; 
        \Phi \vdash G}\rho[\Phi := \ol K] \\ 
      &\hspace{0.3in}|~\forall \ol{K = (K^1, K^2, K^*) : \oCPORT_k}.\\
      &\hspace{0.4in}(\eta_{\overline{K^1}}, \eta_{\overline{K^2}})
      : \oCPORsem{\Gamma; \Phi \vdash F}\Eq_{\rho}[\Phi := \ol K]
      \rightarrow \oCPORsem{\Gamma; \Phi \vdash
        G}\Eq_{\rho}[\Phi := \ol K] \} \\
  \oCPOsem{\Gamma;\Phi \vdash \phi\ol{F}}\rho &=
  (\rho\phi)\,\ol{\oCPOsem{\Gamma;\Phi \vdash
    F}\rho}\\
  \oCPOsem{\Gamma;\Phi \vdash F+G}\rho &=
  \oCPOsem{\Gamma;\Phi \vdash F}\rho +
  \oCPOsem{\Gamma;\Phi \vdash G}\rho\\
  \oCPOsem{\Gamma;\Phi \vdash F \times G}\rho &=
  \oCPOsem{\Gamma;\Phi \vdash F}\rho \times
  \oCPOsem{\Gamma;\Phi \vdash G}\rho\\ 
  \oCPOsem{\Gamma;\Phi,\ol\gamma \vdash (\mu \phi.\lambda
    \ol{\alpha}. H)\ol{G}}\rho &= (\mu
    T^\oCPO_{H,\rho})\ol{\oCPOsem{\Gamma;\Phi,\ol\gamma \vdash G}\rho}\\
    \text{where } T^\oCPO_{H,\rho}\,F & = \lambda
  \ol{A}. \oCPOsem{\Gamma;\ol\gamma,\phi, \ol{\alpha} \vdash
    H}\rho[\phi :=  F][\ol{\alpha := A}]\\
  \text{and } T^\oCPO_{H,\rho}\,\eta &= \lambda
  \ol{A}. \oCPOsem{\Gamma;\ol\gamma,\phi, \ol{\alpha} \vdash
    H}\id_\rho[\phi := \eta][\ol{\alpha := \id_{A}}] \\
  \oCPOsem{\Gamma;\Phi \vdash (\Lan^{\ol{\alpha}}_{\ol{K}} F)\ol{A}}\rho &=
    \{ t : (\textit{Lan}_{\ol{\oCPOsem{\Gamma; \ol\alpha \vdash
          K}\rho[\ol{\alpha := \_}]}} \oCPOsem{\Gamma; \Phi, \ol\alpha
      \vdash F}\rho[\ol{\alpha := \_}]) \ol{\oCPOsem{\Gamma; \Phi
        \vdash A}\rho}\\
    &\hspace{0.3in}|~(t,t) \in \oCPORsem{\Gamma;\Phi \vdash
      (\Lan^{\ol{\alpha}}_{\ol{K}} F)\ol{A}}\Eq_\rho\}\\
\end{align*}
\end{dfn}

@@

@@

If $\Delta = x_1 : F_1,...,x_n : F_n$ is a term context for $\Gamma$
and $\Phi$, then the interpretations $\setsem{\Gamma;\Phi \vdash \Delta}$ and
$\relsem{\Gamma;\Phi \vdash \Delta}$ are defined by
\[\begin{array}{lll}
\setsem{\Gamma;\Phi \vdash \Delta} & = & \setsem{\Gamma;\Phi \vdash
  F_1} \times ... \times \setsem{\Gamma;\Phi \vdash F_n}\\ 
\relsem{\Gamma;\Phi \vdash \Delta} & = & \relsem{\Gamma;\Phi \vdash
  F_1} \times ... \times \relsem{\Gamma;\Phi \vdash F_n}\\ 
\end{array}\]
Every well-formed term $\Gamma;\Phi~|~\Delta \vdash t : F$ then
has, for every $\rho \in \setenv$, set interpretations
$\setsem{\Gamma;\Phi~|~\Delta \vdash t : F}\rho$ as natural
transformations from $\setsem{\Gamma; \Phi \vdash \Delta}\rho$ to
$\setsem{\Gamma; \Phi \vdash F}\rho$, and, for every $\rho \in
\relenv$, relational interpretations $\relsem{\Gamma;\Phi~|~\Delta
  \vdash t : F}\rho$ as natural transformations from
$\relsem{\Gamma; \Phi \vdash \Delta}\rho$ to $\relsem{\Gamma; \Phi
  \vdash F}\rho$. These are given in Definition~\ref{def:set-interp}.

\begin{dfn}\label{def:set-interp}
If $\rho$ is a set (resp., relation) environment and
$\Gamma;\Phi~|~\Delta \vdash t : F$ then
$\setsem{\Gamma;\Phi~|~\Delta \vdash t : F}\rho$ (resp.,
$\relsem{\Gamma;\Phi~|~\Delta \vdash t : F}\rho$) is defined as in
Figure~\ref{fig:term-sem}, where $\mathsf D$ is either $\set$ or
$\rel$ as appropriate.
\end{dfn}

\begin{figure*}
\hspace*{-3.1in}
\resizebox{0.46\linewidth}{!}{
\begin{minipage}[t]{0.5\textwidth}
\dsem{\Gamma;\Phi \,|\, \Delta,x :F \vdash x : F} \rho& = &
\pi_{|\Delta|+1}\\
%\dsem{\Gamma;\emptyset \,|\, \Delta \vdash \lambda x.t : \sigma \to \tau}\rho &
%= & \curry (\dsem{\Gamma;\emptyset \,|\, \Delta, x : \sigma \vdash t :
%  \tau}\rho)\\ 
%\dsem{\Gamma;\emptyset \,|\, \Delta \vdash st: \tau} \rho & = &
%\eval \circ
% \langle \dsem{\Gamma;\emptyset \,|\, \Delta \vdash s: \sigma \to
%  \tau}\rho, \dsem{\Gamma;\emptyset \,|\, \Delta \vdash t: \sigma}\rho
%\rangle\\
\dsem{\Gamma;\emptyset \,|\, \Delta \vdash L_{\overline \alpha} x.t : \Nat^{\overline
    \alpha} \,F \,G}\rho & = &  \curry (\dsem{\Gamma;\overline \alpha
  \,|\, \Delta, x : F \vdash t: G}\rho[\overline{\alpha := \_}])\\
\dsem{\Gamma;\Phi \,|\, \Delta \vdash t_{\overline K} s:
  G [\overline{\alpha := K}]}\rho & = & \eval \circ \langle
  \lambda d.\,(\dsem{\Gamma;\emptyset \,|\, \Delta \vdash t :
  \Nat^{\overline{\alpha}} \,F \,G}\rho\; d)_{\overline{\dsem{\Gamma;\Phi
      \vdash K}\rho}},\\ 
 & & \hspace*{0.5in} \dsem{\Gamma;\Phi \,|\,
    \Delta \vdash s: F [\overline{\alpha := K}]}\rho \rangle\\ 
& & \\
%% \color{red} \mbox{Add rules for } \forall \mbox{ if we include it} & & \\
%\dsem{\Gamma;\Phi \,|\, \Delta,x :F \vdash x : F} \rho& = &
%\pi_{|\Delta|+1}\\
\dsem{\Gamma;\Phi \,|\, \Delta \vdash \bot_F t : F} \rho& = &
!^0_{\dsem{\Gamma;\Phi \vdash F}\rho} \circ
  \dsem{\Gamma;\Phi~|~\Delta \vdash t : \zerot}\rho, \mbox{ where } \\
 & & \hspace*{0.1in} !^0_{\dsem{\Gamma;\Phi \vdash F}\rho}
\mbox{ is the unique morphism from } 0\\
 & & \hspace*{0.1in} \mbox{ to } \dsem{\Gamma;\Phi \vdash F}\rho\\
\dsem{\Gamma;\Phi \,|\, \Delta \vdash \top : \onet}\rho & = &
!^{\dsem{\Gamma;\Phi\vdash \Delta}\rho}_1, \mbox{ where }
!^{\dsem{\Gamma;\Phi\vdash \Delta}\rho}_1\\ 
& & \hspace*{0.1in} \mbox{ is the unique morphism from }
\dsem{\Gamma;\Phi\vdash \Delta}\rho \mbox{ to } 1\vspace*{0.05in}\\ 
\dsem{\Gamma;\Phi \,|\, \Delta \vdash (s,t) : F \times G} \rho& = &
\dsem{\Gamma;\Phi \,|\, \Delta \vdash s: F} \rho\times
\dsem{\Gamma;\Phi \,|\, \Delta \vdash t : G} \rho\\
\dsem{\Gamma;\Phi \,|\, \Delta \vdash \pi_1 t : F} \rho& = &
\pi_1 \circ \dsem{\Gamma;\Phi \,|\, \Delta \vdash t : F \times G}\rho\\
\dsem{\Gamma;\Phi \,|\, \Delta \vdash \pi_2 t : G}\rho & = &
\pi_2 \circ \dsem{\Gamma;\Phi \,|\, \Delta \vdash t : F \times
  G} \rho\\
\dsem{\Gamma;\Phi~|~\Delta \vdash \case{t}{x \mapsto l}{y \mapsto r} :
  K}\rho & = & \eval \circ \langle \curry \,[\dsem{\Gamma;\Phi
    \,|\, \Delta, x : F \vdash l : K}\rho,\\
   & & \hspace*{0.79in} \dsem{\Gamma;\Phi \,|\, \Delta, y
    : G \vdash r : K}\rho],\\
   & &  \hspace*{0.5in} \dsem{\Gamma;\Phi \,|\, \Delta \vdash t :
  F + G} \rho\rangle\vspace*{0.05in}\\   
\dsem{\Gamma;\Phi \,|\, \Delta \vdash \inl \,s: F + G} \rho& = &
\inl \circ \dsem{\Gamma;\Phi \,|\, \Delta \vdash s: F}\rho\\
\dsem{\Gamma;\Phi \,|\, \Delta \vdash \inr \,t: F + G}\rho & = & 
\inr \circ \dsem{\Gamma;\Phi \,|\, \Delta \vdash t : G}\rho\\
\llbracket \Gamma;\emptyset \,|\, \emptyset \vdash \map^{\ol{F},\ol{G}}_H
  : \Nat^\emptyset (\Nat^{\ol\beta,\ol\gamma} F\,G)
& = & \lambda d\, \ol\eta\,\ol{C}.\,
\dsem{\Gamma; \ol{\phi},\ol{\gamma}\vdash H}\id_{\rho[\ol{\gamma:=
      C}]}[\ol{\phi := \lambda \ol B. \eta_{\ol B\, \ol C}}]\\ 
\hspace*{0.5in}
  (\Nat^{\ol{\gamma}}\,H[\ol{\phi :=_{\ol{\beta}} F}]\,H[\ol{\phi
      :=_{\ol{\beta}} G}]) \rrbracket^{\mathsf D} \rho & &\\
%\hspace*{0.79in}  & & \hspace*{0.1in}\id_{\rho[\ol{\gamma :=
%      C}]}[\ol{\phi := \lambda
%    \ol{B}. (\dsem{\Gamma;\ol\alpha\,|\,\Delta \vdash t : 
%    \Nat^{\ol{\beta},\ol{\gamma}}\,F\,G}\rho\,d)_{\ol{B}\,\ol{C}}}]\\
\llbracket \Gamma;\emptyset \,|\, \emptyset \vdash \tin_H :
Nat^{\ol{\beta},\ol{\gamma}} \, H[\phi := (\mu \phi.\lambda {\overline
    \alpha}.H){\overline \beta}][\ol{\alpha := \beta}] & = &
\lambda d\,\ol{B}\,
\ol{C}.\,(\mathit{in}_{{T}^X_{H,\rho[\ol{\gamma :=
        C}]}})_{\ol{B}} \;\;\mbox{ where } X \mbox{ is } \set \mbox{ when } \\ 
\hspace*{0.79in}(\mu \phi.\lambda {\overline \alpha}.H){\overline
  \beta} \rrbracket^{\mathsf{D}} \rho & & \hspace*{0.2in}  
\mathsf{D} = \set \mbox{ and not present when }
\mathsf{D} = \rel\vspace*{0.05in}\\  
\llbracket \Gamma;\emptyset \,|\, \emptyset \vdash
  \fold^F_H : \Nat^\emptyset\;(\Nat^{\ol{\beta}, \ol{\gamma}}\,H[\phi
    :=_{\ol{\beta}} F][\ol{\alpha := \beta}]\,F) & = &  
\lambda d\,\eta\,\ol{B}\,\ol{C}.\,
(\mathit{fold}_{T^X_{H,\rho[\ol{\gamma := C}]}} \, (\lambda
\ol{A}.\,\eta_{\ol{A}\,\ol{C}}))_{\ol{B}}\\ 
\hspace*{0.79in}(\Nat^{{\ol{\beta},\ol{\gamma}} }\,(\mu
  \phi.\lambda \overline \alpha.H)\overline \beta\;F)
\rrbracket^{\mathsf{D}} \rho & & \hspace*{0.2in} \mbox{ where } X
\mbox{ is as above }\vspace*{0.05in} \\
\setsem{\Gamma;\emptyset~|~\emptyset \vdash \int_{\ol K,F}
: \Nat^{\Phi,\ol\alpha}\,F\, (\Lan^{\ol\alpha}_{\ol K}\, F)\ol{K}}\rho
& = & \vspace*{0.05in}\\
\relsem{\Gamma;\emptyset~|~\emptyset \vdash \int_{\ol K,F} :
  \Nat^{\Phi,\ol\alpha}\,F\, (\Lan^{\ol\alpha}_{\ol K}\,
  F)\ol{K}}\rho\,d & = & \vspace*{0.05in} \\
\setsem{\Gamma;\emptyset~|~\Delta \vdash \partial^{G, \ol
      K}_F \eta : Nat^{\Phi, \ol\beta}\, (\Lan^{\ol\alpha}_{\ol K}
  F)\ol \beta\; G}\rho & = &\lambda \ol{N} \, \ol{B}.\,
((\mu_{\ol{N}})_{\ol{B}})|_{\setsem{\Gamma; \Phi, \ol\beta \vdash
    (\Lan^{\ol\alpha}_{\ol K} F)\ol \beta } \rho[\Phi, \ol \beta :=
    \ol N, \ol B]}\\
& & \hspace*{0.2in} \mbox{where } \mu_{\ol{N}} \mbox{ is as described below}
\vspace*{0.05in}\\
\relsem{\Gamma;\emptyset~|~\Delta \vdash \partial^{G, \ol
      K}_F \eta : Nat^{\Phi, \ol\beta}\, (\Lan^{\ol\alpha}_{\ol K}
    F)\ol \beta\; G}\rho & = & \\
\caption{Term semantics}\label{fig:term-sem} 
\end{minipage}}\vspace*{-0.05in}

\new{In the fourth-to-last clause of Definition~\ref{def:set-interp},...}

\new{In the third-to-last clause of Definition~\ref{def:set-interp},...}

In the next-to-last clause of Definition~\ref{def:set-interp},
  \[\mu_{\ol{N}} : \Lan_{\ol{\setsem{\Gamma; \ol\alpha \vdash K}\rho
  {[\ol{\alpha := \_} ]}}} \setsem{\Gamma; \Phi, \ol\alpha \vdash
    F}\rho {[\Phi, \ol \alpha := \ol N, \_ ]} \to \setsem{\Gamma;
    \Phi, \ol\beta \vdash G} \rho[\Phi, \ol \beta := \ol N, \_]\]
  is the unique natural transformation such that
\begin{equation}\label{eq:defn-of-mu-sub-N}
    (\setsem{\Gamma; \emptyset \,|\, \Delta \vdash \eta : \Nat^{\Phi,
    \ol\alpha} \, F \,\, G[\ol{\beta := K}]}\rho \, d)_{\ol{N}} =
\mu_{\ol{N}} \, \ol{\setsem{\Gamma; \ol\alpha \vdash K}
  \rho[\ol{\alpha := \_}]} \circ \epsilon_{\ol{N}}
  \end{equation}
Here, $\epsilon_{\ol{N}}$ is the natural transformation associated
with the left Kan extension
\[\begin{tikzcd} [column sep = large, row sep = large]
\set^{|\ol \alpha|}
\ar[rr, "\setsem{\Gamma; \Phi, \ol\alpha \vdash F}\rho {[\Phi, \ol
      \alpha := \ol N, \_ ]}"{name=F}] 
\ar[dr, "\ol{\setsem{\Gamma; \ol\alpha \vdash K}\rho {[\ol{\alpha := \_} ]}}"{left}] && \set \\ 
&\set^{|\ol K|} \ar[ur, "\Lan_{\ol{\setsem{\Gamma; \ol\alpha \vdash K}\rho {[\ol{\alpha := \_} ]}}}
\setsem{\Gamma; \Phi, \ol\alpha \vdash F}\rho {[\Phi, \ol \alpha :=
    \ol N, \_ ]}"{right}]
\ar[d, Rightarrow, shorten <= 2mm, shorten >= 1mm, 
          from=F, to=2-2,  "\epsilon_{\ol{N}}"{}]
\end{tikzcd}\]

\new{In the final last clause of Definition~\ref{def:set-interp},...}

\vspace*{1in}

If $t$ is closed, i.e., if $\emptyset; \emptyset~|~\emptyset \vdash t
: F$, then we write $\dsem{\vdash t : F}$ instead of $\dsem{\emptyset;
  \emptyset~|~\emptyset \vdash t : F}$.  The return type for the
semantic fold is $\dsem{\Gamma;\ol\beta,\ol\gamma \vdash
  F}\rho[\ol{\gamma := C}][\ol{\beta := B}]$.  This interpretation
gives $\dsem{\Gamma;\emptyset \,|\, \Delta \vdash \lambda x.t : F \to
  G}\rho = \curry (\dsem{\Gamma;\emptyset \,|\, \Delta, x : F \vdash t
  : G}\rho)$ and $\dsem{\Gamma;\emptyset \,|\, \Delta \vdash st: G}
\rho = \eval \circ \langle \dsem{\Gamma;\emptyset \,|\, \Delta \vdash
  s: F \to G}\rho, \dsem{\Gamma;\emptyset \,|\, \Delta \vdash t:
  F}\rho \rangle$, so it specializes to the standard interpretations
for System F terms.  To see that the object and relational
interpretations of $\Gamma;\emptyset~|~\emptyset \vdash \int_{\ol K,F}
: \Nat^{\Phi,\ol\alpha}\,F\, (\Lan^{\ol\alpha}_{\ol K}\, F)\ol{K}$ and
$\Gamma;\emptyset~|~\Delta \vdash \partial^{G, \ol K}_F \eta :
\Nat^{\Phi, \ol\beta}\, (\Lan^{\ol\alpha}_{\ol K} F)\ol \beta\; G$ are
well-defined is significantly more involved, as we see in the next
four subsections.

\subsection{$\setsem{\Gamma;\emptyset~|~\emptyset \vdash \int_{\ol K,F}
: \Nat^{\Phi,\ol\alpha}\,F\, (\Lan^{\ol\alpha}_{\ol K}\, F)\ol{K}}$ is
  well-defined}

\subsection{$\relsem{\Gamma;\emptyset~|~\emptyset \vdash \int_{\ol K,F}
: \Nat^{\Phi,\ol\alpha}\,F\, (\Lan^{\ol\alpha}_{\ol K}\, F)\ol{K}}$ is
  well-defined}

\subsection{$\setsem{\Gamma;\emptyset~|~\Delta \vdash \partial^{G, \ol
      K}_F \eta : Nat^{\Phi, \ol\beta}\, (\Lan^{\ol\alpha}_{\ol K}
    F)\ol \beta\; G}$ is well-defined} We must show that, for every
$\rho : \setenv$ and $d : \setsem{\Gamma; \emptyset \vdash \Delta}
\rho$, \[\setsem{\Gamma; \emptyset \,|\, \Delta \vdash \partial^{G,
    \ol K}_F \eta : \Nat^{\Phi, \ol\beta}\, (\Lan^{\ol\alpha}_{\ol K}
  F)\ol \beta\; G} \rho \, d\] has the correct type and satisfies the
condition required for it to be in $\setsem{\Gamma; \emptyset \vdash
  \Nat^{\Phi, \ol\beta}\, (\Lan^{\ol\alpha}_{\ol K} F)\ol \beta\; G}
\rho$. For the former we observe that each component
$(\mu_{\ol{N}})_{\ol{B}}$ of $\mu_{\ol{N}}$ has type
  \[\Lan_{\ol{\setsem{\Gamma; \ol\alpha \vdash K}\rho
  {[\ol{\alpha := B} ]}}} \setsem{\Gamma; \Phi, \ol\alpha \vdash F}\rho
        {[\Phi, \ol \alpha := \ol N, \ol B]} \to \setsem{\Gamma; \Phi,
          \ol\beta \vdash G} \rho[\Phi, \ol \beta := \ol N, \ol B]\]
        Each component of $(\mu_{\ol{N}})_{\ol{B}} |_{\setsem{\Gamma;
            \Phi, \ol\beta \vdash (\Lan^{\ol\alpha}_{\ol K} F)\ol
            \beta } \rho[\Phi, \ol \beta := \ol N, \ol B]}$ therefore
        has the correct type to be the component at $\ol N$ and $\ol
        B$ of a natural transformation in $\setsem{\Gamma; \emptyset
          \vdash \Nat^{\Phi, \ol\beta}\, (\Lan^{\ol\alpha}_{\ol K}
          F)\ol \beta\; G} \rho$.

The next two lemmas show that $ \lambda \ol{N} \,
\ol{B}. (\mu_{\ol{N}})_{\ol{B}} |_{\setsem{\Gamma; \Phi, \ol\beta
    \vdash (\Lan^{\ol\alpha}_{\ol K} F)\ol \beta } \rho[\ol{\Phi,
      \beta := N, B}]}$ is natural in both $\ol{N}$ and $\ol{B}$ and
that it satisfies the condition required for it to be in
$\setsem{\Gamma; \emptyset \vdash \Nat^{\Phi, \ol\beta}\,
  (\Lan^{\ol\alpha}_{\ol K} F)\ol \beta\; G} \rho$, namely, if $\Phi =
{\phi_1,...,\phi_n}$ and $\phi_i$ has arity $k_i$ for $i = 1,...,n$,
then 
\begin{equation}\label{eq:P2}
    \begin{split}
      &\forall \ol{N = (N^1, N^2, N^*) : RT_{k_i}}, \forall \ol{B =
        (B^1, B^2, B^*) :
        RT_0}.\\ &((\mu_{\ol{N^1}})_{\ol{B^1}}|_{\setsem{\Gamma; \Phi,
          \ol\beta \vdash (\Lan^{\ol\beta}_{\ol K} F)\ol \beta }
        \rho[\Phi, \ol \beta := \ol{N^1}, \ol{B^1}]} ,
      (\mu_{\ol{N^2}})_{\ol{B^2}}|_{\setsem{\Gamma; \Phi, \ol\beta
          \vdash (\Lan^{\ol\beta}_{\ol K} F)\ol \beta } \rho[\Phi, \ol
          \beta := \ol{N^2}, \ol{B^2}]} )\\ &: \relsem{\Gamma; \Phi,
        \ol\beta \vdash (\Lan^{\ol\beta}_{\ol K} F)\ol \beta
      }\Eq_{\rho}[\Phi, \ol \beta := \ol N, \ol B] \rightarrow
      \relsem{\Gamma; \Phi, \ol\beta \vdash G}\Eq_{\rho}[\Phi, \ol
        \beta := \ol N, \ol B]
    \end{split}
    \end{equation}

\begin{lemma}
For every $\rho : \setenv$, $d : \setsem{\Gamma; \emptyset \vdash
  \Delta} \rho$, $\ol{N = (N^1, N^2, N^*) : RT_{k_i}}$, $\ol{B = (B^1,
  B^2, B^*) : RT_0}$, and $(t_1, t_2) \in \relsem{\Gamma; \Phi,
  \ol\beta \vdash (\Lan^{\ol\beta}_{\ol K} F)\ol \beta
}\Eq_{\rho}[\Phi, \ol \beta := \ol N, \ol B]$, it holds that
\[((\mu_{\ol{N^1}})_{\ol{B^1}} \,\, t_1, (\mu_{\ol{N^2}})_{\ol{B^2}} \,\, t_2) \in
\relsem{\Gamma; \Phi, \ol\beta \vdash G}\Eq_{\rho}[\Phi, \ol \beta :=
  \ol N, \ol B]\]
\end{lemma}

\begin{proof}
Unfolding the definition of relational interpretation for $\Lan$-types
in the hypothesis gives that $(t_1, t_2) : \setsem{\Gamma;\Phi,
  \ol\beta \vdash (\Lan^{\ol{\alpha}}_{\ol{K}} F)\ol{\beta}}\rho[\Phi,
  \ol \beta := \ol{N^1}, \ol{B^1}] \,\times\, \setsem{\Gamma;\Phi,
  \ol\beta \vdash (\Lan^{\ol{\alpha}}_{\ol{K}} F)\ol{\beta}}\rho[\Phi,
  \ol \beta := \ol{N^2}, \ol{B^2}]$ and there exist $\ol{Z : \set_0},
t_1' : \setsem{\Gamma; \Phi, \ol\alpha \vdash F} \rho[\ol{\Phi :=
    N^1}][\ol{\alpha := Z}],\, t_2' : \setsem{\Gamma; \Phi, \ol\alpha
  \vdash F}\rho[\ol{\Phi := N^2 }][\ol{\alpha := Z}]$,\\ $\ol{f :
  \relsem{\Gamma; \ol{\alpha} \vdash K}\Eq_\rho[\ol{\alpha := \Eq_Z}]
  \to B}$ \, such that $\iota_{\ol{Z},\ol{\pi_1 f}}t_1' = t_1,\,
\kappa_{\ol{Z}, \ol{\pi_2 f}}t_2' = t_2,$ and $(t_1',t_2') \in
\\ \relsem{\Gamma; \Phi, \ol\alpha \vdash F}\Eq_\rho[\ol{\Phi :=
    N}][\ol{\alpha := \Eq_Z}].$ Inlining $t_1 =
\iota_{\ol{Z},\ol{\pi_1 f}}t_1'$ and $t_2 = \kappa_{\ol{Z}, \ol{\pi_2
    f}}t_2'$ in in the goal in the statement of the lemma gives us the
following rewritten goal to prove:
\begin{equation}\label{eq:new-goal}
((\mu_{\ol{N^1}})_{\ol{B^1}} \,\, \iota_{\ol{Z},\ol{\pi_1 f}}\, t_1' ,
  (\mu_{\ol{N^2}})_{\ol{B^2}} \,\,  \kappa_{\ol{Z}, \ol{\pi_2 f}}\, t_2'  )
\in \relsem{\Gamma; \Phi, \ol\beta \vdash
G}\Eq_{\rho}[\ol{\Phi, \beta := N, B}]
\end{equation}

To this end, consider the commuting diagram\\

\begin{tikzcd}[column sep = large, row sep = large]
    & \setsem{\Gamma; \Phi, \ol\alpha \vdash F}\rho[\Phi, \ol \alpha
    := \ol{N^1}, \ol Z]
    \ar[dr, "\;\;\;\;\;\iota_{\ol{Z}, \ol{\pi_1 f}}"{right}] 
    \ar[dl, "\iota'_{\ol Z, 
                \ol{\id_{\setsem{\Gamma; \ol\alpha \vdash K}\rho[\ol{\alpha := Z}]}}}\;\;\;\;\;\;\;\;\;\;\;\;"{left}] 
        & \\
    LKZ % (\Lan_{\setsem{\Gamma; \ol\alpha \vdash K}\rho{[\ol{\alpha := \_}]}}
    %               \setsem{\Gamma; \Phi, \ol\alpha \vdash F}
    %                 \rho{[\ol{\Phi, \alpha := N^1, \_}]}) \ol{\setsem{\Gamma; \ol\alpha \vdash K}\rho[\ol{\alpha := Z}]}
    \ar[rr, dashed, "(\Lan_{\ol{\setsem{\Gamma; \ol\alpha \vdash K}\rho{[\ol{\alpha := \_}]}}}
                  \setsem{\Gamma; \Phi, \ol\alpha \vdash F}
                    \rho{[\Phi, \ol \alpha := \ol{N^1}, \_]}) \ol{\pi_1 f}"{below}]
       & & LB
  \end{tikzcd}
